---
title: "Homework 4"
subtitle: STA 138 | Camden Possinger | Winter 2022
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
wd <- "/home/cam/Documents/STA 138/Homework"
knitr::opts_knit$set(root.dir = wd)
```

```{r, include = FALSE}
library(magrittr)
library(kableExtra)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggpubr)
```

## 1.   For the model $\log(\frac{\pi}{1-\pi}) = \alpha + \beta_1x_1 + \beta_2x_2$


```{r}

angina <- read.csv("angina.csv")
angina %<>% select(y,myofam,age)

angina$myofam %>% as.factor 

```


### (a) What are the estimated parameters for this model?

```{r}

no_interact <- glm(y ~ myofam + age, data = angina, family = "binomial")
est_params <- data.frame("Parameter" = c("Alpha","Beta 1","Beta 2"),
                         "Estimated_Values" = no_interact$coefficients)

est_params %>% kbl

```


### (b) Interpret the parameters. From this model, what would you estimate to be the probability of angina for a patient with age 50 and with a family history of myocardial infarction?


$\hat{\alpha}$ is the estimated log-odds ratio when a patient does not have a family history of myocardial infarction and has 0 age  

$\hat{\beta_1}$ is the estimated log-odds ratio of angina between when a patient does have a family history of myocardial infarction and when a patient does not keeping age constant.  

$\hat{\beta_2}$ is the estimated log-odds ratio of angina based on a 1 year change in age keeping everything else constant. 


```{r}

est_odds <- no_interact %>% predict(newdata = list("myofam" = "yes", "age" = 50)) %>% unname 

est_prob <- est_odds %>% plogis

print(est_prob)

```
Here the estimated probability of this individual who has a family history of myocardial infarction and is aged 50 is about 0.78, which is fairly high.


### (c) Plot both the fitted log-odds and fitted probability of angina for patients as a function of age and myofam status.

```{r}
max_age <- angina$age %>% max
min_age <- angina$age %>% min

age <- min_age:max_age


fitted_odds_1 <- no_interact %>% predict(newdata = data.frame("age" = age, "myofam" = rep("yes",age %>% length))) %>% unname 
fitted_odds_0 <- no_interact %>% predict(newdata = data.frame("age" = age, "myofam" = rep("no",age %>% length))) %>% unname 

fitted_probs_1 <- fitted_odds_1 %>% plogis
fitted_probs_0 <- fitted_odds_0 %>% plogis

plot_data <- data.frame(age,fitted_odds_1,fitted_odds_0,fitted_probs_1,fitted_odds_0)

fitted_odds_plot <- ggplot()+geom_line(data = plot_data, aes(x = age, y = fitted_odds_1,color = "yes"))+
  geom_line(data = plot_data, aes(x = age, y = fitted_odds_0,color = "no"))+
  xlab("Age")+
  ylab("Fitted Log-Odds")

fitted_probs_plot <-  ggplot()+geom_line(data = plot_data, aes(x = age, y = fitted_probs_1,color = "yes"))+
  geom_line(data = plot_data, aes(x = age, y = fitted_probs_0,color = "no"))+
  xlab("Age")+
  ylab("Fitted Probabilities")


ggarrange(fitted_odds_plot, fitted_probs_plot, nrow = 1)

```

### (d) Under this model, estimate the odds ratio of angina for a patient if they did have a family history of myocardial infarction vs. the same patient if they didn’t have such a history. Does it depend on age?

```{r}

est_odds_ratios <- data.frame( "Parameter" = c("Alpha","Beta 1","Beta 2"), 
                         "Estimated_Odds_Ratios" = no_interact$coefficients %>% exp)

est_odds_ratios %>% kbl


```

In this case the estimated log odds ratios for angina for a patient shows that as the patient ages the odds of angina increases by about 9% while if a patient has a family history of myocardial infarction the odds of angina increases by about 742%. Angina does depend on age, but mostly depends on a family history of myocardial infarction. Here the odds ratio for family history of myocardial infarction does not depend on age.

$\frac{exp(-3.1481617 + (0.0887068)x_2)}{exp(-5.2783645 + (0.0887068)x_2)}$ does not depend on $x_2$



## 2.   For the model $\log(\frac{\pi}{1-\pi}) = \alpha + \beta_1x_1 + \beta_2x_2 + \beta_3 x_1 \cdot x_2$


### (a) What are the estimated parameters for this model?

```{r}

interact <- glm(y ~ myofam + age + myofam:age, data = angina, family = "binomial")
est_params <- data.frame("Parameter" = c("Alpha","Beta 1","Beta 2","Beta 3"),
                         "Estimated_Values" = interact$coefficients)

est_params %>% kbl

```


### (b) Interpret the parameters. From this model, what would you estimate to be the probability of angina for a patient with age 50 and with a family history of myocardial infarction?

$\hat{\alpha}$ is the estimated log odds-ratio of angina when both age = 0 and a patient does not have a family history of myocardial infarction.

$\hat{\beta_1}$ is the estimated log odds-ratio of angina between if a patient has a family history of myocardial infarction or not when age = 0. 

$\hat{\beta_2}$ is the estimated log odds-ratio of angina between a one year difference in a patient's age when a patient does not have a family history of myocardial infarction. 

$\hat{\beta_3}$ quantifies the interaction effect between age and family history of myocardial infarction on the presence of angina. 


```{r}

est_odds <- interact %>% predict(newdata = list("myofam" = "yes", "age" = 50, "myofamyes:age" = 50)) %>% unname 

est_prob <- est_odds %>% plogis

print(est_prob)

```
Here the estimated probability of this individual who has a family history of myocardial infarction and is aged 50 is about 0.80, which is still high and doesn't change much from the the model in question 1 


### (c) Plot both the fitted log-odds and fitted probability of angina for patients as a function of age and myofam status.

```{r}

fitted_odds_1 <- interact %>% predict(newdata = data.frame("age" = age, "myofam" = rep("yes",age %>% length))) %>% unname 
fitted_odds_0 <- interact %>% predict(newdata = data.frame("age" = age, "myofam" = rep("no",age %>% length))) %>% unname 

fitted_probs_1 <- fitted_odds_1 %>% plogis
fitted_probs_0 <- fitted_odds_0 %>% plogis

plot_data <- data.frame(age,fitted_odds_1,fitted_odds_0,fitted_probs_1,fitted_odds_0)

fitted_odds_plot <- ggplot()+geom_line(data = plot_data, aes(x = age, y = fitted_odds_1,color = "yes"))+
  geom_line(data = plot_data, aes(x = age, y = fitted_odds_0,color = "no"))+
  xlab("Age")+
  ylab("Fitted Log-Odds")

fitted_probs_plot <-  ggplot()+geom_line(data = plot_data, aes(x = age, y = fitted_probs_1,color = "yes"))+
  geom_line(data = plot_data, aes(x = age, y = fitted_probs_0,color = "no"))+
  xlab("Age")+
  ylab("Fitted Probabilities")


ggarrange(fitted_odds_plot, fitted_probs_plot, nrow = 1)




```

### (d) Under this model, estimate the odds ratio of angina for a patient if they did have a family history of myocardial infarction vs. the same patient if they didn’t have such a history. Does it depend on age?

```{r}

est_odds_ratios <- data.frame("Parameter" = c("Alpha","Beta 1","Beta 2","Beta 3"),      
                         "Estimated_Odds_Ratios" = interact$coefficients %>% exp)

est_odds_ratios %>% kbl

interact_sum <- interact %>% summary
p_values <- data.frame("Parameter" = c("Alpha","Beta 1","Beta 2","Beta 3"),      
                         "p-values" =  interact_sum$coefficients[,4] )
p_values %>% kbl
```

Here the odds-ratio depends on age since the interaction term is included. 

$\frac{exp(-7.1879268 + (0.1715519)x_2)}{exp(-4.0470314 + (0.0655033)x_2)}$ depends on $x_2$
