---
title: "Homework 5"
subtitle: STA 138 | Camden Possinger | Winter 2022
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
wd <- "/home/cam/Documents/STA 138/Homework"
knitr::opts_knit$set(root.dir = wd)
```

```{r, include = FALSE}
library(magrittr)
library(kableExtra)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggpubr)
```

```{r}

largework <- read.csv("largework.csv")

#largework$gender %<>% as.factor 
#largework$marriage %<>% as.factor 
#largework$y %<>% as.factor 

```


## 1. Randomly subset about half of your data as a training set, and using this, report the best model according to forward stepwise selection with AIC that you obtain.

```{r}

set.seed(502)

largework_len <- largework %>% nrow

train_set_ind <- sample(largework_len,largework_len / 2) 

train_set <- largework[train_set_ind,]
test_set <- largework[-train_set_ind,]


best_model <- step(glm(y~1,family = "binomial",data = train_set),
                   scope = ~gender*age*marriage*min*chol*sysbp*height,
                   direction = "forward",
                   trace = 0)

print(best_model$formula)

```

## 2.  Using the model obtained above with the testing set, test for evidence of nonzero coefficients, using α = 0.1; Interpret your results.

```{r}
best_model_test_sum <- glm(y ~ chol + sysbp + gender + height:chol + height:sysbp + sysbp:gender, "binomial", test_set) %>% summary

best_model_train_sum <- best_model %>% summary

print("Test Set: ")
best_model_test_sum$coefficients[,c(1,4)]


print("Training Set: ")
best_model_train_sum$coefficients[,c(1,4)]



```
In this model using the test set there are no variables that are significant using $\alpha = 0.1$ which using Bonferroni correction would be $\alpha = 0.1/7$. This is interesting because when we use the training set there are variables that are significant. 


